---
title: "Ocean pH - solution to exercise"
author: "your name"
date: "the date"
format: html
editor: visual
---

A consequence of the increase in the atmospheric concentration of carbon dioxide (CO2) that has occurred over the last 250 years is that the oceans are absorbing more CO2 and as a consequence, becoming more acidic

You have been supplied with a data set taken from a figure published by the IPCC, which shows ocean pH vs time for the last 40 years or so.

The data are in the file `ipcc_AR6_WGI_Figure_5_20-pH.csv`. Note

Your task is to write a script that uses the exact same sequence of steps as was used to analyse plant growth rate to determine whether there is evidence for any linear decline in pH over this perid, and , if so, what the rate of decline is: by how much does ocean pH change each year?

Once you have finished your analysis, report your conclusions in pain English.

You can use the plant growth rate script as a guide, and also the session notes. Note that the data in the csv file do not begin until line 6. Prior to that there is some meta data that tells you where the data comes from.

## Load packages
```{r}
library(tidyverse)
library(here)
library(ggfortify)
library(cowplot)
```

## Load data
```{r}
filepath <- here("data","ipcc_AR6_WGI_Figure_5_20-pH.csv")
pH_data <- read_csv(filepath, skip = 6)
glimpse(pH_data)
```
## Plot the data
```{r}
pH_data |>
  ggplot(aes(x = year, y = pH)) +
  geom_point() +
  labs(x="Year",
       y="Ocean pH") +
  theme_cowplot()
```

There is a plausibly linear variation of pH with year since the 1980s. The variance of the data is roughly constant throughout this period.


## Make a linear model of the variation with time of ocean since the mid 1980s.

```{r}
model_pH<-lm(pH ~ year, data = pH_data)
```


## Check assumptions

```{r}
autoplot(model_pH, smooth.colour=NA) + theme_cowplot()
```


```{r}
anova(model_pH)
```
The F value is very large, the p value is tiny (much less than 0.05) so we rekjct the null hypothesis that pH is not changing with year.

## Effect size
```{r}
summary(model_pH)
```
The ocean pH has been reducing at the rate of 0.00158 per year (and thus 0.0158 per decade) throughout this period. 

## Report the model
We find that ocean pH has been reducing at the rate of 0.00158 per year since the 1980s (p<0.001). A linear regression model accounts for 71% of the variability in the data.


